---
title: "Take Home Exercise 2"
description: |
  Take Home Exercise 2: Focusing on Airbnb and how their expansion has impacted our economy. Using Spatial Point Patterns Analysis of Airbnb Listing in Singapore. 
author:
  - name: Sarah Chin
    url: linkedin.com/in/sarahchin99/
date: 09-14-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 1. Overview 

Airbnb has expanded their services over 34,000 cities across 191 countries. However, Singapore is still one of the global cities that has yet to legalise short-term rentals offered by platforms such as Airbnb. Despite Singapore's disregard of using Airbnb, there are still tools and datasets about Singapore that allows people to explore how Airbnb are used in the cities. 


# 2. Installing and Loading the packages 

``` {r echo=TRUE, eval=TRUE}
packages = c('maptools', 'sf', 'raster','spatstat', 'tmap', 'onemapsgapi')
for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p,character.only = T)
}
```

# 3. Section A: Airbnb Distribution in 2019

In this section, we need to investigate if the distribution of Airbnb listings are affected by location factors such as near to existing hotels, MRT services and tourist attractions. 

Before we can analyse these points, we need to import and clean our data. Firstly, we import the Airbnb data using _st_read()_ of sf package and transform the coordinate system to 3414. 

``` {r echo=TRUE, eval=TRUE}
airbnb <- read.csv("Airbnb_listing_30062019/30062019.csv")
```

We also want to extract the number and locations of hotels and tourist attractions in Singapore to see how this competition affects the Airbnb listings. 

``` {r echo=TRUE, eval=TRUE}
hotels <- read.csv("OneMap_Data/hotels.csv")
tourism <- read.csv("OneMap_Data/tourism.csv")
```

Extracting the data for MRT stations are also important for analysis. As the data for MRT station is in _shp_ file format, we will use the following code to extract the data. 

``` {r echo=TRUE, eval=TRUE}
mrt <- st_read(dsn = "TrainStation", 
                layer = "MRTLRTStnPtt")
```

Since the MRT dataset is a _shp_ file, it can be plotted immediately. 

``` {r}
tmap_mode("view")
tm_shape(mrt) + 
  tm_dots()
```

Since the Airbnb, hotels and tourists datasets that have been imported are in _.csv_ format, we would need to convert them to _sf_ for further analysis. Additionally, we need to change the coordinate system to _3414_, the coordinate system of Singapore. As all of the data provided for latitude and longitude are in decimal degree format, we will assume that the data is in *wgs84* Geographic Coordinate System. 

``` {r echo=TRUE, eval=TRUE}
airbnb_sf <- st_as_sf(airbnb, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)

hotels_sf <- st_as_sf(hotels, 
                       coords = c("Lng", "Lat"),
                       crs=4326) %>%
  st_transform(crs = 3414)

tourism_sf <- st_as_sf(tourism, 
                       coords = c("Lng", "Lat"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

Let's plot to review the datasets that have been provided. This is the Airbnb map using _airbnb_sf_.

``` {r echo=TRUE, eval=TRUE}
tmap_mode("view")
tm_shape(airbnb_sf) + 
  tm_dots(alpha = 0.4, 
          col = "blue", 
          size = 0.05)
```

Here is the hotels map using _hotels_sf_.

``` {r echo=TRUE, eval=TRUE}
tm_shape(hotels_sf) + 
  tm_dots(alpha = 0.4, 
          col = "red", 
          size = 0.05)
```

Here are the tourist attractions available in Singapore, using _tourism_sf_.

``` {r echo=TRUE, eval=TRUE}
tm_shape(tourism_sf) + 
  tm_dots(alpha = 0.4, 
          col = "purple", 
          size = 0.05)
```

As we can see from the above results for _tourism_sf_, there is a coordinate that is not within Singapore. This means that this point (Longitude and Latitude) could possibly be N/A. We can verify this by searching for any missing values. 

``` {r echo=TRUE, eval=TRUE}
sum(is.na(tourism_sf$LATITUDE))
```

From the results, we can tell that there is one N/A result in the column _"LATITUDE"_ under the _tourism_sf_ dataset. We shall remove that N/A value to concentrate our findings on Singapore. 

``` {r echo=TRUE, eval=TRUE}
tourism_sf <- tourism_sf[!is.na(tourism_sf$LATITUDE),]
```

After the N/A row has been removed, we can plot the graph again to see if there's an improvement. 

``` {r echo=TRUE, eval=TRUE}
tm_shape(tourism_sf) + 
  tm_dots(alpha = 0.4, 
          col = "purple", 
          size = 0.05)
```

After cleaning the _tourism_sf_ dataset, we can finally put the 3 datasets together to see if there are any correlation between the datasets. The Airbnb dataset are highlighted in blue, the hotels dataset are highlighted in red and the tourism dataset highlighted in purple. Here's all the datasets together: 

``` {r echo=TRUE, eval=TRUE}
tmap_mode("view")
tm_shape(airbnb_sf) + 
  tm_dots(alpha = 0.4, 
          col = "blue", 
          size = 0.05) +
tm_shape(hotels_sf) + 
  tm_dots(alpha = 0.4, 
          col = "red", 
          size = 0.05) +
tm_shape(tourism_sf) + 
  tm_dots(alpha = 0.4, 
          col = "purple", 
          size = 0.05) +
tm_shape(mrt) +
  tm_dots(alpha = 0.4, 
          col = "green", 
          size = 0.05)
```

From the above plotted map, we can tell that the Airbnb facilities have been spread widely over Singapore, covering places that even the hotels are not available in. On the other hand, majority of the hotels are located in the central district of Singapore with the exception of some hotels such as RM Hotel on the far west and Changi hotels in the east. However, the location of the hotels can be related to the tourism locations. As seen above, the locations of most of the tourist attractions are within the central district of Singapore as well. In order to capitalise and profit from tourists, hotels would locate themselves nearer to the tourist attractions as tourists would prefer to be nearer to these attractions. For the MRT lines, we can see that it correlates closely with the Airbnb locations, especially in the Northern region of Singapore. This could be because the Airbnb hosts believe that being near an MRT station would be more attractive to tourists for easy access to transport. 

Now that we have plotted our graph, we can start the geospatial data wrangling process. 

## Geospatial Data Wrangling

One of the objectives in this task is to derive the kernel density maps of the Airbnb listings, hotels, MRT services and tourist attractions. In order to analyse any of the data that we have plotted so far, we would need to further clean the data with the following steps. 

### Step 1: Converting sf data frames to sp's Spatial class

As the _airbnb_sf_, _hotels_sf_ and _tourism_sf_ are all in sf data frame, we would need to first convert them into Spatial class. 

``` {r echo=TRUE, eval=TRUE}
airbnb_spatial <- as_Spatial(airbnb_sf)
hotels_spatial <- as_Spatial(hotels_sf)
tourism_spatial <- as_Spatial(tourism_sf)
mrt_spatial <- as_Spatial(mrt)
```

``` {r echo=FALSE, eval=TRUE}
airbnb_spatial
```

``` {r echo=FALSE, eval=TRUE}
hotels_spatial
```

``` {r echo=FALSE, eval=TRUE}
tourism_spatial
```

``` {r echo=FALSE, eval=TRUE}
mrt_spatial
```

### Step 2: Converting Spatial class into sp format 

As *spatstat* requires the data in *_ppp_* format and there is no direct way to convert *Spatial* class into *_ppp_*, we need to first convert the data into *Spatial* object. 

``` {r echo=TRUE, eval=TRUE}
airbnb_sp <- as(airbnb_spatial, "SpatialPoints")
hotels_sp <- as(hotels_spatial, "SpatialPoints")
tourism_sp <- as(tourism_spatial, "SpatialPoints")
mrt_sp <- as(mrt_spatial, "SpatialPoints")
```

``` {r echo=FALSE, eval=TRUE}
airbnb_sp
```

``` {r echo=FALSE, eval=TRUE}
hotels_sp
```

``` {r echo=FALSE, eval=TRUE}
tourism_sp
```

``` {r echo=FALSE, eval=TRUE}
mrt_sp
```

### Step 3: Converting sp format into spatstat's ppp format 

Now that the datasets are in _sp_ format, we can finally convert the datasets into _ppp_ object format. 

``` {r echo=TRUE, eval=TRUE}
airbnb_ppp <- as(airbnb_sp, "ppp")
hotels_ppp <- as(hotels_sp, "ppp")
tourism_ppp <- as(tourism_sp, "ppp")
mrt_ppp <- as(mrt_sp, "ppp")
```

``` {r echo=FALSE, eval=TRUE}
airbnb_ppp
```

``` {r echo=FALSE, eval=TRUE}
hotels_ppp 
```

``` {r echo=FALSE, eval=TRUE}
tourism_ppp
```

``` {r echo=FALSE, eval=TRUE}
mrt_ppp
```

### Handling duplicated points 

Before we can proceed, we need to check to see if the data contains any duplicated points. We can do so by using the _any(duplicated())_ function. 

``` {r echo=TRUE, eval=TRUE}
any(duplicated(airbnb_ppp))
any(duplicated(hotels_ppp))
any(duplicated(tourism_ppp))
any(duplicated(mrt_ppp))
```

From the above results, we can tell that 3 datasets, excluding _mrt_pp_, contain duplicated points. Therefore, we need to properly handle them before moving on. We can use jittering, which is a solution that adds a small perturbation to the duplicated points, ensuring that the points do not occupy the same space. We can use this solution by using the following code: 

``` {r echo=TRUE, eval=TRUE}
airbnb_ppp_jit <- rjitter(airbnb_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)

hotels_ppp_jit <- rjitter(hotels_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)

tourism_ppp_jit <- rjitter(tourism_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

After running the code above, let's check to see if there are any duplicated points left. 

``` {r echo=TRUE, eval=TRUE}
any(duplicated(airbnb_ppp_jit))
any(duplicated(hotels_ppp_jit))
any(duplicated(tourism_ppp_jit))
```

### Creating owin object 

In order to properly analyse the data within Singapore, we need to create an object owin to represent the Singapore boundary as a polygonal region. We will use the _MPSZ_SUBZONE_WEB_PL_ dataset that was used in Hands on exercise 4. 

``` {r echo=TRUE, eval=TRUE}
mpsz_sf <- st_read(dsn = "data", 
                layer = "MP14_SUBZONE_WEB_PL")

mpsz_spatial <- as_Spatial(mpsz_sf)

mpsz_sp <- as(mpsz_spatial, "SpatialPolygons")

mpsz_owin <- as(mpsz_sp, "owin")
```

Next, let's plot _mpsz_owin_ to see the subzones and the outline of Singapore. 

``` {r}
plot(mpsz_owin)
```

### Combining the owin object with the points object 

Lastly, we extract the individual datasets' points that are located within Singapore using this code: 

``` {r echo=TRUE, eval=TRUE}
airbnbSG_ppp <- airbnb_ppp[mpsz_owin]
hotelsSG_ppp <- hotels_ppp[mpsz_owin]
tourismSG_ppp <- tourism_ppp[mpsz_owin]
mrtSG_ppp <- mrt_ppp[mpsz_owin]
```

**Airbnb**

``` {r echo=TRUE, eval=TRUE}
plot(airbnbSG_ppp)
```

**Hotels**

``` {r echo=TRUE, eval=TRUE}
plot(hotelsSG_ppp)
```

**Tourist Attractions**

``` {r echo=TRUE, eval=TRUE}
plot(tourismSG_ppp)
```

*MRT Stations*

``` {r echo=TRUE, eval=TRUE}
plot(mrtSG_ppp)
```

## First-Order Spatial Point Patterns Analysis 

Now that we have found the _ppp_ values of all the datasets, we can proceed to calculate the kernel density estimation (KDE) layer to visualise and explore the intensity of point processes. 

Using the code chunk below, we will be able to produce and plot the kernel density of all the datasets. 

**Airbnb**

``` {r echo=TRUE, eval=TRUE}
kde_airbnb_bw <- density(airbnbSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 

plot(kde_airbnb_bw)
```

As we can see from the plot above, the output range is too small for us to make any analysis. We will need to convert the unit of measure from meter to kilometer. 

``` {r echo=FALSE, eval=TRUE}
airbnbSG_ppp.km <- rescale(airbnbSG_ppp, 1000, "km")
```

Now, we will be able to re-run the code and plot the KDE map. 

``` {r echo=TRUE, eval=TRUE}
kde_airbnbSG.bw <- density(airbnbSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_airbnbSG.bw, main = "Airbnb bw.ppl")
```

Now, we will do the same KDE plotting for the rest of the datasets: Hotels, Tourist Attractions and MRT stations. 

**Hotels**

``` {r echo=TRUE, eval=TRUE}
kde_hotels_bw <- density(hotelsSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 

plot(kde_hotels_bw)
```

Similar to the Airbnb dataset, we shall convert the unit of measurement and plot the density map again. 

``` {r echo=TRUE, eval=TRUE}
hotelsSG_ppp.km <- rescale(hotelsSG_ppp, 1000, "km")

kde_hotelsSG.bw <- density(hotelsSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_hotelsSG.bw, main = "Hotels bw.ppl")
```

**Tourist Attractions**

``` {r echo=TRUE, eval=TRUE}
kde_tourism_bw <- density(tourismSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 

plot(kde_tourism_bw)
```

``` {r echo=TRUE, eval=TRUE}
tourismSG_ppp.km <- rescale(tourismSG_ppp, 1000, "km")

kde_tourismSG.bw <- density(tourismSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_tourismSG.bw, main = "Tourist Attractions bw.ppl")
```

**MRT Stations**

``` {r echo=TRUE, eval=TRUE}
kde_mrt_bw <- density(mrtSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 

plot(kde_mrt_bw)
```

``` {r echo=TRUE, eval=TRUE}
mrtSG_ppp.km <- rescale(mrtSG_ppp, 1000, "km")

kde_mrtSG.bw <- density(mrtSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_mrtSG.bw, main = "MRT Stations bw.ppl")
```

Now that the First Order Analysis has been completed, we can move on to the Second Order Spatial Point Patterns Analysis. However, we would need to analyze the spatial points using G-Function before we can do so. In order to do that, we must first conduct the Nearest Neighbour Analysis

## Nearest Neighbour Analysis 

For Nearest Neighbour Analysis, we will be using the Clark-Evans test of aggregation for a spatial point pattern analysis. 

The test hypotheses will adjust according to the four different factors: Airbnb, Hotels, Tourist Attractions and MRT Station. Here are the hypotheses: 

Ho = The distribution of each factor are randomly distributed.

H1 = The distribution of each factor are not randomly distributed. 

For this hypothesis, we will use the 95% confident interval. 

__Airbnb__

``` {r echo=TRUE, eval=TRUE}
clarkevans.test(airbnbSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

__Hotels__

``` {r echo=TRUE, eval=TRUE}
clarkevans.test(hotelsSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

__Tourist Attractions__

``` {r echo=TRUE, eval=TRUE}
clarkevans.test(tourismSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

__MRT Stations__

``` {r echo=TRUE, eval=TRUE}
clarkevans.test(mrtSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

As seen above, all the tests have resulted in _alternative hypothesis: clustered (R < 1)_. This means that the point pattern of all the datasets have resulted in the alternative hypothesis, the clustered point pattern. Now that we have performed the Clark-Evans test, we can proceed with the Second-Order Spatial Point Patterns Analysis. 

## Second-Order Spatial Point Patterns Analysis

For the Second-Order Spatial Point Patterns Analysis, we will use the K-function to estimate the distribution of the different factors. 

First, let's take a look at the Airbnb dataset. We will use _Fest()_ to compute the F-function of the dataset. 

``` {r echo=TRUE, eval=TRUE}
F_airbnb = Fest(airbnb_ppp)
plot(F_airbnb)
```

As seen in the graph above, the estimated line, _K(iso)_, is much higher than the actual line, _K(pois)_. Now we will now perform the Complete Spatial Randomness Test. We will have the following hypothesis: 

Ho = The distribution of Airbnb locations are randomly distributed. 

H1 = The distribution of Airbnb locations are not randomly distributed. 

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001. 

``` {r echo=TRUE, eval=TRUE}
F_airbnb.csr <- envelope(airbnb_ppp, Fest, nsim = 999)
```

``` {r echo=TRUE, eval=TRUE}
plot(F_airbnb.csr)
```





















