---
title: "In Class Exercise 5"
description: |
  A short description of the post.
author:
  - name: Sarah Chin
    url: linkedin.com/in/sarahchin99/
date: 09-13-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Installing and Loading the R package

+ maptools is for creating spatial objects
+ sf is for importing geospatial data
+ raster is for creating raster object
+ spatstat is for performing spatial point pattern analysis
+ tmap is for plotting thematic maps 
+ tidyverse is for data handling 

``` {r echo=TRUE, eval=TRUE}
packages <- c('maptools', 'sf', 'raster', 'spatstat', 'tmap', 'tidyverse', 'plotly', 'ggthemes')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
library(p, character.only = T)
}
```


## Importing Geospatial data 

Importing shapefile using _st_read()_ of *sf* package. The output object is in tibble sf object class. 

``` {r echo=TRUE, eval=TRUE}
mpsz_sf <- st_read(dsn = "data/shapefile",
                   layer = "MP14_SUBZONE_WEB_PL") # st_read is for geospatial data, not for excel 
# for excel, we have readxl
```
Projection is in SVY21. 


## Importing Aspatial data

*read_rds()* of **readr** package is used instead of *readRDS()* of base R is used. This is because output of *read_RDS* is in tibble object. 

``` {r echo=TRUE, eval=TRUE}
CHAS <- read_rds("data/rds/CHAS.rds")
childcare <- read_rds("data/rds/childcare.rds")
```


Note that there are some data issue in childcare data frame because 'Lat' and 'Lng' should be in numeric data type. 

* The coordinate fields seem to be in decimal degree. Hence, WGS referencing system is assumed. 

## Converting the aspatial data frame into sf objects

``` {r echo=TRUE, eval=TRUE}
CHAS_sf <- st_as_sf(CHAS,
                    coords = c("X_COORDINATE",
                               "Y_COORDINATE"), # can refer to the excel itself for the coords
                    crs = 3414)
```

Note: *st_as_sf* accept coordinates in character data type. 

``` {r echo=TRUE, eval=TRUE}
childcare_sf <- st_as_sf(childcare,
                    coords = c("Lng",
                               "Lat"), 
                    crs = 4326) %>% # because it is in decimal 
  st_transform(crs = 3414)
```

If you want to change the 'Lat' and 'Lng' in the original dataset to numeric data type, it is also acceptable. 

# Plot map to review 

``` {r echo=TRUE, eval=TRUE}
tmap_mode("view")
tm_shape(childcare_sf) + 
  tm_dots(alpha = 0.4, # alpha sets the transparency
          col = "blue", 
          size = 0.05)
```

You can also view more than one data in a single map.

``` {r echo=TRUE, eval=TRUE}
tmap_mode("view")
tm_shape(childcare_sf) + 
  tm_dots(alpha = 0.4, 
          col = "blue", 
          size = 0.05) +
tm_shape(CHAS_sf) + 
  tm_dots(alpha = 0.4, 
          col = "red", 
          size = 0.05)
```

## Geospatial Data Wrangling

### Converting from sf to Spatial dataframe classes

as_Spatial() of sf package. 

``` {r echo=TRUE, eval=TRUE}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <- as_Spatial(mpsz_sf)
```

### Converting Spatial data frame into Spatial objects 

as.SpatialPoint() of as.SpatialPolygon() of maptools package.

``` {r echo=TRUE, eval=TRUE}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS, "SpatialPoints")
mpsz_sp <- as(mpsz, "SpatialPolygons")
```

Note that it is no longer a data frame, it is a formal class. Compare it to _childcare_ and you realise that the data has been dropped from _childcare_sp_. 

### Converting from Spatial Objects into ppp objects

Dropping all the projection information. All that would be left is the coordinates. 

* Using as.ppp() of maptools package to do the conversion

``` {r echo=TRUE, eval=TRUE}
childcare_ppp <- as(childcare_sp, "ppp") 
CHAS_ppp <- as(CHAS_sp, "ppp")

```

### Removing duplicate points using jitter 

``` {r echo=TRUE, eval=TRUE}
childcare_ppp_jit <- rjitter(childcare_ppp,
                             retry = TRUE, 
                             nsim = 1, # number of simulations
                             drop = TRUE)
any(duplicated(childcare_ppp_jit)) # checks if there are any duplicated in the jit version of childcare
```

``` {r echo=TRUE, eval=TRUE}
CHAS_ppp_jit <- rjitter(CHAS_ppp,
                             retry = TRUE, 
                             nsim = 1, 
                             drop = TRUE)
any(duplicated(CHAS_ppp_jit)) 
```

Now we want to plot it in a graph. However, we cannot use _tmap_ like how we did before since it is in *ppp* format. We have to use the _plot()_ function instead. 

### Extracting Punggol Planning Area 

Specifying which of the data we want to use.

``` {r echo=TRUE, eval=TRUE}
pg <- mpsz[mpsz@data$PLN_AREA_N=="PUNGGOL",]
```

### Converting SpatialPolygonDataFrame into SpatialPologons object

``` {r echo=TRUE, eval=TRUE}
pg_sp <- as(pg, "SpatialPolygons")
```

### Converting SpatialPolygons into owin object

``` {r echo=TRUE, eval=TRUE}
pg_owin <- as(pg_sp, "owin")
```

### Extracting spatial points window owin

``` {r echo=TRUE, eval=TRUE}
childcare_pg <- childcare_ppp_jit[pg_owin]
CHAS_pg <- CHAS_ppp_jit[pg_owin]
```

```{r echo=TRUE, eval=TRUE}
plot(childcare_pg)
```

## L-Function

``` {r echo=TRUE, eval=TRUE}
L_childcare <- envelope(childcare_pg,
                        Lest, 
                        nsim = 99, 
                        rank = 1, 
                        global = TRUE)
```

After the iterations are done, it should go through this chunk in order to get an interactive version of the map. 

```{r echo=TRUE, eval=FALSE}
title <- "Pairwise Distance: L function"

Lcsr_df <- as.data.frame(L_childcare)

colour=c("#0D657D","#ee770d","#D3D3D3")
csr_plot <- ggplot(Lcsr_df, aes(r, obs-r))+
  # plot observed value
  geom_line(colour=c("#4d4d4d"))+
  geom_line(aes(r,theo-r), colour="red", linetype = "dashed")+
  # plot simulation envelopes
  geom_ribbon(aes(ymin=lo-r,ymax=hi-r),alpha=0.1, colour=c("#91bfdb")) +
  xlab("Distance r (m)") +
  ylab("L(r)-r") +
  geom_rug(data=Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,], sides="b", colour=colour[1])  +
  geom_rug(data=Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,], sides="b", colour=colour[2]) +
  geom_rug(data=Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,], sides="b", color=colour[3]) +
  theme_tufte()+
  ggtitle(title)

text1<-"Significant clustering"
text2<-"Significant segregation"
text3<-"Not significant clustering/segregation"

# the below conditional statement is required to ensure that the labels (text1/2/3) are assigned to the correct traces
if (nrow(Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,])==0){ 
  if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text3, traces = 4) %>%
      rangeslider() 
  }else if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      rangeslider() 
  }else {
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider() 
  }
} else if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){
  if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      rangeslider() 
  } else{
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider()
  }
} else{
  ggplotly(csr_plot, dynamicTicks=T) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    rangeslider()
  }
```

